Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2022-03-22 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Add consensus standards and core samples sections. Reorganize.

Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite currents. This is in line with what other hybrid sources produce. Typical "good" currents are between 5-10uA. The source typically produces best current at around 5uLs-1 CO2. Literature values give 1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and flows in at least one of these papers doesn't make sense.

@@

## Reference materials
The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.

Measured

Table: summary of reference materials

figure5: difference plot

Assessment of data quality.

@@

To further test the method and a potential application, we measured 15 shell and 
coral macrofossils from a subtropical coastal marine sediment core using our gas 
method and compared to the same samples measured using the traditional 
graphite method. 
We found the HGIS and graphite measurements to be in good agreement, with no 
significant bias (figure 6). Mean per-sample error was 1 pMC for HGIS measurements, 
compared with 0.3 pMC for graphite.
The accuracy of age-depth models is often driven by environmental factors and 
sparse data, rather than the precision of individual dates [3]. The lower time and 
cost of the hybrid method allows analysis of more samples. This may produce more 
accurate age models, despite reduced precision of individual measurements.

### Measurements

Table of measurements (supplementary?)

One HGIS sample vial was noted to be at atmospheric pressure when only XX mL CO2 should have been produced. The HGIS measurement for this sample was XX% higher than the measurement as graphite. We suspect that this sample was improperly sealed and eliminated it from the analysis as an outlier. SO2 story?

Figure6: bland-altman of HGIS vs graphite.

Summary of agreement

@@

@@

## Quantification of method

## Cost vs quality discussion

@@

# Conclusion

We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source. This system produces results that agree with graphite measurements to better than 1% from XXX mg of carbonate. Sample preparation takes about 10 min. of person time and 6h of automated processing and reaction time, and 20 min. of measurement time per sample. The simplicity of this approach... With lower cost and time per sample, the lower precision of the method may be balanced by the ability to measure more samples from a population, ie. triplicate measurement or more core depths. This may lead to more accurate results compared to a more precise graphite measurement where reproducibility is not as well constrained.

>>>>

Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite currents. This is generally comparable with other hybrid sources REFS. Typical currents for this study were between 4-10uA???. The source typically produces best current at around 5uLs-1 CO2. Literature values give 1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and flows in at least one of these papers doesn't make sense.

@@

## Reference materials

Precision and accuracy of the method were assessed by measuring carbonate standards with consensus values from graphite measurements. Summary data for these samples are listed in table 1, and measurement data for all samples is in supplementary information. Measurement, data reduction, and blank correction was performed as described above.

Overall, reference materials run with the hybrid gas method were in good agreement with consensus values. Mean difference from consenus for all measurements was -0.04 pMC, and the standard deviation of the difference was 1.10 pMC. Mean propagated measurement error was 1.13 pMC for these samples. The hybrid method has no bias in agregate. Method error is larger than graphite, but calculated per-sample error is similar to standard deviation of replicate measurements. A summary of accuracy and precision of these measurements is in Table 1.

The difference beween reference samples and graphite consensus does not vary systematically by fraction modern, indicating that the method has no significant bias by fraction modern (Figure 6).

The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.

@@

To further test the method and a potential application, we measured 15 shell and 
coral macrofossils from a subtropical coastal marine sediment core using our gas 
method and compared to the same samples measured using the traditional 
graphite method.

We found the HGIS and graphite measurements to be in good agreement, with no 
significant bias (figure 6). Mean difference between hybrid and graphite measurements was -0.35 pMC and the standard deviation of this difference was 0.85 pMC. Mean per-sample error was 1.02 pMC for HGIS measurements, compared with 0.18 pMC for graphite.

One HGIS sample vial was noted to be at atmospheric pressure when the 16mg sample should have produced only one half atmosphere of CO2. The HGIS measurement for this sample was 5.3 pMC higher than the measurement as graphite. We suspect that this sample was improperly sealed and eliminated it from the analysis as an outlier.

@@

Age models were developed for both the traditional graphite measurements of the core carbonates and the HGIS measurements (Figure 7). We used the Bchron Baesian radiochronology package to develop these models. The age-depth curves produced by the two datasets are very similar, with the largest differences being in the uncertainty of the individual dates.

Mean and sd of difference of calibrated ages?

@@

We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source. This system produces results that agree with graphite measurements to better than 1% from 8-20 mg of carbonate. Sample preparation takes about 10 min. of person time and 6h of automated processing and reaction time, and 20 min. of measurement time per sample. The simplicity of this approach... With lower cost and time per sample, the lower precision of the method may be balanced by the ability to measure more samples from a population, ie. triplicate measurement or more core depths. This may lead to more accurate results compared to a more precise graphite measurement where reproducibility is not as well constrained.

@@
